
# Multiple Regression

This chapter will show you how to add two, three, and even more numeric explanatory variables to a linear model.

## Fitting a MLR model

In terms of the R code, fitting a multiple linear regression model is easy: simply add variables to the model formula you specify in the `lm()` command.

In a parallel slopes model, we had two explanatory variables: one was numeric and one was categorical. Here, we will allow both explanatory variables to be numeric.

____________

```{r}
library(ggplot2)
load("./Data/mario_kart.RData")
str(mario_kart)
```

The dataset `mario_kart` is already loaded in your workspace.

* Fit a multiple linear regression model for total price as a function of the duration of the auction and the starting price.

```{r}
# Fit the model using duration and startPr
mod <- lm(totalPr ~ duration + startPr, data = mario_kart)
mod
```

_______________

## Tiling the plane

One method for visualizing a multiple linear regression model is to create a [heatmap](https://en.wikipedia.org/wiki/Heat_map) of the fitted values in the plane defined by the two explanatory variables. This heatmap will illustrate how the model output changes over different combinations of the explanatory variables.

This is a multistep process:

* First, create a grid of the possible pairs of values of the explanatory variables. The grid should be over the actual range of the data present in each variable. We've done this for you and stored the result as a data frame called `grid`.

```{r}
grid <- expand.grid(duration = seq(1, 10, by = 1), startPr = seq(0.01, 69.95, by = 0.01))
```

* Use `augment()` with the `newdata` argument to find the $\hat{y}$'s corresponding to the values in `grid`.

* Add these to the `data_space` plot by using the fill aesthetic and `geom_tile()`.

```{r, fig.width = 6}
data_space <- ggplot(data = mario_kart, 
                     aes(x = duration, y = startPr)) + 
  geom_point(aes(color = totalPr)) + 
  theme_bw()
data_space
```


_________________

### Exercise{-}

The model object `mod` is already in your workspace.

* Use `augment()` to create a `data.frame` that contains the values the model outputs for each row of `grid`.

```{r}
# add predictions to grid
price_hats <- broom::augment(mod, newdata = grid)
```

* Use `geom_tile` to illustrate these predicted values over the `data_space` plot. Use the `fill` aesthetic and set `alpha = 0.5`.

```{r, fig.width = 6}
# tile the plane
data_space + 
   geom_tile(data = price_hats, 
             aes(fill = .fitted), alpha = 0.5)
```

_____________

## Models in 3D

An alternative way to visualize a multiple regression model with two numeric explanatory variables is as a plane in three dimensions. This is possible in R using the `plotly` package.

We have created three objects that you will need:

* `x`: a vector of unique values of `duration`
* `y`: a vector of unique values of `startPr`
* `plane`: a matrix of the fitted values across all combinations of `x` and `y`

Much like `ggplot()`, the `plot_ly()` function will allow you to create a plot object with variables mapped to `x`, `y`, and `z` aesthetics. The `add_markers()` function is similar to `geom_point()` in that it allows you to add points to your 3D plot.

Note that `plot_ly` uses the pipe (`%>%)` operator to chain commands together.

____________

### Exercise {-}

* Run the `plot_ly` command to draw 3D scatterplot for `totalPr` as a function of duration and `startPr` by mapping the `z` variable to the response and the `x` and `y` variables to the explanatory variables. Duration should be on the x-axis and starting price should be on the y-axis.

```{r, fig.width = 8, fig.height = 8}
library(plotly)
# draw the 3D scatterplot
p <- plot_ly(data = mario_kart, z = ~totalPr, x = ~duration, y = ~startPr, opacity = 0.6) %>%
  add_markers() 
p
```


* Use `add_surface()` to draw a plane through the cloud of points by setting `z = ~plane`.  See [wikipedia](https://en.wikipedia.org/wiki/Outer_product) for the definition of an outer product.  In what follows, we will use the R function `outer()` to compute the values of `plane`.

$$\bf{u \otimes v} = \bf{uv^T}$$

```{r, fig.width = 8, fig.height = 8}
summary(mod)$coef
x <- seq(1, 10, length = 70)
y <- seq(0.010, 59.950, length = 70)
plane <- outer(x, y, function(a, b){summary(mod)$coef[1,1] + 
    summary(mod)$coef[2,1]*a + summary(mod)$coef[3,1]*b})
# draw the plane
p %>%
  add_surface(x = ~x, y = ~y, z = ~plane, showscale = FALSE)
```

___________________

### Coefficient magnitude{-}

The coefficients from our model for the total auction price of MarioKarts as a function of auction duration and starting price are shown below.

```{r}
mod
```

A colleague claims that these results imply that the duration of the auction is a more important determinant of final price than starting price, because the coefficient is larger. This interpretation is false because:

__________

* The coefficient on duration is negative.

* Smaller coefficients are more important.

* **The coefficients have different units (dollars per day and dollars per dollar, respectively) and so they are not directly comparable.**

* The intercept coefficient is much bigger, so it is the most important one.

_________________

### Practicing interpretation{-}

Fit a multiple regression model for the total auction price of an item in the `mario_kart` data set as a function of the starting price and the duration of the auction. Compute the coefficients and choose the correct interpretation of the duration variable.

_____________

* **For each additional day the auction lasts, the expected final price declines by $1.51, after controlling for starting price.**

* For each additional dollar of starting price, the expected final price increases by $0.23, after controlling for the duration of the auction.

* The duration of the auction is a more important determinant of final price than starting price, because the coefficient is larger.

* The average auction lasts 51 days.

____________________

## Visualizing parallel planes

By including the duration, starting price, and condition variables in our model, we now have two explanatory variables and one categorical variable. Our model now takes the geometric form of two parallel planes!

The first plane corresponds to the model output when the condition of the item is `new`, while the second plane corresponds to the model output when the condition of the item is `used`. The planes have the same slopes along both the duration and starting price axes---it is the z-intercept that is different.

Once again we have stored the `x` and `y` vectors for you. Since we now have two planes, there are matrix objects `plane0` and `plane1` stored for you as well.

```{r}
modI <- lm(totalPr ~ duration + startPr + cond, data = mario_kart)
summary(modI)$coef
plane0 <- outer(x, y, function(a, b){53.3447530 -0.6559841*a + 
                                      0.1981653*b})
plane1 <- outer(x, y, function(a, b){53.3447530 -0.6559841*a +
                                      0.1981653*b - 8.9493214})
```

___________

### Exercise {-}

* Use `plot_ly` to draw 3D scatterplot for `totalPr` as a function of `duration`, `startPr`, and `cond` by mapping the `z` variable to the response and the `x` and `y` variables to the explanatory variables. Duration should be on the x-axis and starting price should be on the y-axis. Use color to represent `cond`.

```{r, fig.width = 8, fig.height = 8}
# draw the 3D scatterplot
p <- plot_ly(data = mario_kart, z = ~totalPr, x = ~duration, y = ~startPr, opacity = 0.6) %>%
  add_markers(color = ~cond) 
p
```

* Use `add_surface()` (twice) to draw two planes through the cloud of points, one for new MarioKarts and another for used ones. Use the objects `plane0` and `plane1`.

```{r, fig.width = 8, fig.height = 8}
# draw two planes
p %>%
  add_surface(x = ~x, y = ~y, z = ~plane0, showscale = FALSE) %>%
  add_surface(x = ~x, y = ~y, z = ~plane1, showscale = FALSE)
```

____________

### Parallel plane interpretation{-}

The coefficients from our parallel planes model is shown below.

```{r}
modI
```

Choose the right interpretation of $\beta_3$ (the coefficient on `condUsed`):

________

* **The expected premium for new (relative to used) MarioKarts is $8.95, after controlling for the duration and starting price of the auction.**

* The expected premium for used (relative to new) MarioKarts is $8.95, after controlling for the duration and starting price of the auction.

* For each additional day the auction lasts, the expected final price declines by $8.95, after controlling for starting price and condition.

___________

### Interpretation of coefficient in a big model {-}

This time we have thrown even more variables into our model, including the number of bids in each auction (`nBids`) and the number of wheels. Unfortunately this makes a full visualization of our model impossible, but we can still interpret the coefficients.


```{r}
modJ <- lm(totalPr ~ duration + startPr + cond + wheels + nBids, 
    data = mario_kart)
modJ
```

Choose the correct interpretation of the coefficient on the number of wheels:

________

* The average number of wheels is 6.72.

* Each additional wheel costs exactly $6.72.

* Each additional wheel is associated with an increase in the expected auction price of $6.72.

* **Each additional wheel is associated with an increase in the expected auction price of $6.72, after controlling for auction duration, starting price, number of bids, and the condition of the item.**

_______________